WORCS: A Workflow for Open Reproducible Code in Science

### 2020-11-17 paper/reviews_ds.Rmd
more extensive discussion of licensing

Dear Dr. Editor,

thank you very much for taking the time to consider our manuscript for publication at _`r rmarkdown::metadata$journal`_.
In the following we address your and each reviewers' concerns point-by-point.

@@

To address the Reviewer's comment, we have set up an evaluation questionnaire, and sent it to the attendees of our previous workshops on WORCS. At the time of writing, however, this has not yielded actionable points of improvement.

We do have a strong history of addressing user comments submitted through the GitHub issues page.
WORCS adheres to the Core Infrastructure Initiative "best practices" (CII Best Practices),
which requires that
the channels through which users can provide feedback are clearly indicated,
as they are [here](https://github.com/cjvanlissa/worcs#contributing-and-contact-information).
It is also evident from the [resolved tickets on GitHub](https://github.com/cjvanlissa/worcs/issues?q=is%3Aissue+is%3Aclosed) that we have been very responsive in actively incorporating user feedback - as also required by the CII Best Practices.

We are committed to continue evaluating user experience, both actively through evaluation forms, and passively through the GitHub issues pages.
We hope that this commitment, along with the evidence of our prior responsiveness to user feedback presented above, assuage the Reviewer's concerns.

Finally, we wish to point out that the Editor has indicated that 'resource papers do not need to have a full-blown user evaluation, but we expect "sound evidence of its (potential for) reuse"'. We therefore elaborate on the efforts we have made to ensure that WORCS is reusable, followed by evidence of its current adoption rates.

@@

Second, there is some evidence to demonstrate that WORCS is already being adopted (as of 31-10-2020):

1. [The preprint](https://psyarxiv.com/k4wde/) has been downloaded 568 times, since being published on 31-05-2020
1. The `worcs` R-package has been downloaded 2474 times from CRAN, since being published on 18-05-2020
1. The GitHub project has been forked by users unknown to us 4 times, and starred (= watched) 32 times
1. The video tutorial for worcs has been viewed 362 times, and liked 18 times (https://www.youtube.com/watch?v=uzjpN_yFeUU)
1. WORCS is used to teach reproducibility at Utrecht University
1. The lead author has given invited lectures on WORCS at the Erasmus University Rotterdam, Eindhoven University of Technology, and Max Planck Institute, Berlin
1. The lead author has applied `worcs` in several research projects:
	* https://github.com/cjvanlissa/storm_meta_father_mother
	* https://github.com/cjvanlissa/veni_sysrev
	* https://github.com/cjvanlissa/placebo_adhd
	* https://github.com/cjvanlissa/schumpe
	* https://github.com/cjvanlissa/coping_covid

@@

Response: In addressing this comment, we respectfully refer to our response to the preceding comment ("There is no user evaluation", R1), and again emphasize the Editor's comment that "resource papers do not need to have a full-blown user evaluation". We would also like to clarify that the goal of the present manuscript is to define the problem WORCS intends to address (specifically; meeting the TOP-guidelines for Open Science), introduce the tools WORCS uses to address this problem, explain how WORCS uses these tools, and provide a step-by-step workflow. It is not clear what "claims" are made that are unsupported by evidence. In the revision, we have attempted to further clarify this goal from the paper onset: REF.

We do take the suggestion of performing a user evaluation very seriously, but in our mind, this is a next step. We have built WORCS based on established tools and best practices, and we have gathered preliminary positive feedback from the community through workshops and teaching.

@@

Response: We understand that our writing comes across as contradictory,
and this apparent contradiction arises from a lack of explicit definitions of the terms "replication" and "reproducibility" in the original manuscript.
We are aware of the oft-lamented lack of consistent definitions for these terms (e.g., Patil, Peng, & Leek, 2019; Plesser, 2018).
In the revision, we have attempted to address this lack of clarity by including explicit definitions for these terms:

@@

After making these definitions explicit, it becomes clear that the Reviewer's definition of replication as "a necessary steps towards reproducibility" is not consistent with the definitions of the present paper.

We further wish to clarify that the fact that we address guidelines 1-7, but not guideline 8 ("replication") is consistent with the intended purpose of the TOP-guidelines.
Specifically, the first seven guidelines are intended to ensure the "reproducibility of the reported results based on the originating data" [direct quote from @nosekPromotingOpenResearch2015a]. The eighth criterion, by contrast, is "not formally a transparency standard for authors, [but] addresses journal guidelines for consideration of independent replications for publication" (again a direct quote).
To avoid further confusion or controversy, we now address this distinction between the first 7 guidelines and the final 8th explicitly in the revised manuscript:

> These guidelines describe eight standards for open science,
the first seven of which "account for reproducibility of the reported results
based on the originating data, and for sharing sufficient information to conduct an independent 
replication" [@nosekPromotingOpenResearch2015a].
In this context, 
reproducibility can be defined as "re-performing the same analysis with the same code and the same data using a different analyst",
and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].
These guidelines are: 1) Comprehensive
citation of literature, data, materials, and methods; 2) sharing data, 3)
sharing the code required to reproduce analyses, 4) sharing new research
materials, and 5) sharing details of the design and analysis; 6)
pre-registration of studies before data collection, and 7) pre-registration of
the analysis plan prior to analysis.
The eighth criterion is "not formally a transparency standard for authors", but "addresses 
journal guidelines for consideration of independent replications for publication".
WORCS defines the goals of open science in terms of the first seven TOP-guidelines,
and the workflow and its `R` implementation are designed to facilitate compliance therewith.

@@

Response: We apologize; it appears that there were numerous errors in the parsing of the reference list after rendering the manuscript to HTML - the preferred format of Data Science. The reference is indeed to the R-package. In the revised manuscript, we have repaired the references so that it is clear that this is a software package.

@@

@@

\Assignment{BV, AB, AL, CJ}
\WorkInProgress
\Done

@@

We agree completely that it is important to make open data findable beyond just providing it freely.
This is achieved by linking the GitHub repository to an OSF.io project page;
on OSF, the project receives a DOI.
It is also possible to give a separate DOI to specific resources, such as data files.
We agree that this was not emphasized enough in the manuscript, and have rewritten the revision to clarify.

It is a pity that the Reviewer does not provide arguments to support a preference for Zenodo or FigShare over GitHub, because if there is a strong case to be made, we would be keen to include this in the paper.
Without such arguments, as far as we are concerned, these repositories are largely interchangeable.

We prefer storing the data on GitHub because they are then bundled with the
documentation (e.g., codebook), analysis code, and manuscript.
Thus, readers have access to **all** materials required for replication in one repository.
GitHub additionally offers user interaction with the materials through
forking, cloning, opening issues, and sending pull requests -
all useful tools for replication and collaboration.

It is further worth noting that [GitHub is compatible with Zenodo](https://guides.github.com/activities/citable-code/);
users can document GitHub repositories in Zenodo and generate a DOI.
This is redundant with Step 15 of the workflow, however:
Connecting the GitHub repository to an OSF project accomplishes effectively the same.
We do acknowledge, however, 
that this part of the workflow was not clearly documented in the previous version of the manuscript, 
and that the use of DOIs warrants further clarification.

In the revision, we have attempted to address this comment in two ways: First, by
clarifying Step 15 in the WORCS workflow, specifically emphasizing the importance of generating a DOI. 
Second, by discussing the use of DOIs in the paper.

The approach we advocate is to make all of a project's resources available in a 
Step 15 of the workflow, "Connect your ‘OSF’ project page to the ‘Git’ remote repository",
was intended to meet this goal.

Digital object identifiers are the de-factor standard for making unique references to data sets. In line with the reviewer's suggestion,
we suggest to use external providers to provide such DOIs while github is still the primary data storage. TODO!!!

BV: Can you please argue for why GitHub is equivalent to Zenodo for this purpose?
AL: Can you address the reference to FAIR principles in relation to GitHub?

AB: "on second thought, it’s probably worth mentioning to the reviewer that we know these options but defend the OSF-centric approach for our purposes. My impression is that the main challenge are the different expectations and solutions across disciplines that are currently evolving largely in parallel. If we acknowledge that there are alternatives, this may please the reviewers while we still can defend R + OSF for the behavioral and social sciences. Other disciplines may have different foci, e.g., python + jupyter + docker etc... and no single manuscript will be able to cover everything."

\Assignment{BV, AB, AL, CJ}
\WorkInProgress
\Hard

@@

> For users of the `worcs` `R` package, we highly recommend referring to
[the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html),
which describes in more detail which steps are automated by the `R` package
and which actions must be taken by the user.
The vignette is also updated with future developments of the package.

@@

> We provide a conceptual outline of the WORCS procedure below, based on WORCS
Version 0.1.5.
The conceptual workflow recommends actions to take 
in the different phases of a research project, in order to work openly and transparently
in compliance with the TOP-guidelines and best practices for open science.
Note that, although the steps are numbered for reference purposes,
we acknowledge that the process of conducting research is not always linear.

@@

Response: We wish to emphasize that it was certainly unintended to claim that everyone should us R for open science. To address this comment, we have done two things: First, throughout the paper, we have attempted to clarify the distinction between the platform independent conceptual workflow "WORCS", and its implementation for R users, `worcs`. Second, throughout the paper, we encourage the implementation of WORCS in other languages. Most illustrative of these changes is the start of the paragraph on **The R implementation of WORCS**:

@@

\Assignment{ALL}
\WorkInProgress
\Medium

@@

Anna-Lena?

Response: 
F: Dit heeft weer met DOIs te maken. Ik denk dat als we het gebruik van OSF de default van worcs maken (die mijn opmerking R1C7a) dat we hier een sterk punt kunnen maken mbt findability. 
A: Eigenlijk hetzelfde antwoord als bij Findability: dit wordt heel eenvouding als we de OSF route voor data storage de default maken. 
I: Het feit dat we alleen gebruik maken van open-source software maakt alle projecten die worcs gebruiken nogal interoperable. 
R: Dit is wel een belangrijkt punt: Kunnen we opnemen in de workflow dat worcs-projecten by default een CC-BY licentie krijgen?

\Assignment{AL, BV, CJ}
\WorkInProgress
\Medium

@@

Response: WORCS does not provide a service for de-identification,
and we consider this to be outside of the scope of the present paper.
If the Reviewer feels differently, we are open to being convinced, however!
Of course, since `worcs` is open source software, anybody is welcome to contribute this functionality in exchange for coauthorship, in line with the contribution guidelines [outlined here](https://github.com/cjvanlissa/worcs#contributing-and-contact-information).
The current team lacks relevant expertise to make this contribution.
<!--What would this entail? Is it within scope?-->

@@

To address the Reviewer's comment, we have made two changes: First, we now explicitly state, in the paper and the Workflow vignette, that the steps are not necessarily linear:  *"Note that, although the steps are numbered for reference purposes, we acknowledge that the process of conducting research is not always linear."*. Second, while it was always possible to manually add a manuscript or preregistration at a later point in time after creating a new project in RStudio - we have added more functions to increase the non-linear flexibility of the `worcs` package: `add_manuscript()` and `add_preregistration()`.

@@

Response: Respectfully, we are not entirely sure what the criticism is here.
First, as addressed in our response to the previous comment, there seems to be a misunderstanding about the scope of the paper. To clarify, we again state that our paper introduces a workflow for open science, based on objective criteria set out in the TOP-guidelines. Regarding novelty: Aside from the work in progress of our co-authors - which is focused on strict computational reproducibility and is compatible with the present workflow - this is one of the first efforts to translate the requirements of open science into a step-by-step workflow. When you search Google Scholar for "open science workflow", the first hit that actually describes a workflow is the present manuscript.

@@

> Git tracks changes to files on a line-by-line basis. The user can sture these changes by making a "commit": a 
snapshot of the version controlled files.
Each "commit" can contain as many changes as desired;
for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
Each commit is further tagged with a message, describing the changes.
The `R` implementation of `worcs` contains a convenience function,
`git_update("your commit message")`,
which creates a new commit for all changes since the previous commit,
labels it with "your commit message", and pushes these changes to the remote repository.
It is, effectively, a Git "quick save" command.

@@

Response: We thank the Reviewer for acknowledging the idea to be interesting.
Whether it will be adopted by researchers remains to be seen after publication.
There is some prelimiary evidence that the workflow is being adopted (as of 31-10-2020):

1. The preprint has been downloaded 561 times, since being published on 31-05-2020
1. The `worcs` R-package has been downloaded 2401 times from CRAN, since being published on 18-05-2020
1. The GitHub project has been forked by users unknown to us 4 times, and starred (= watched) 32 times
1. The video tutorial for worcs has been viewed 362 times, and liked 18 times (https://www.youtube.com/watch?v=uzjpN_yFeUU)
1. WORCS is used to teach reproducibility at Utrecht University
1. The lead author has given invited lectures on WORCS at the Erasmus University Rotterdam, Eindhoven University of Technology, and Max Planck Institute, Berlin
1. The lead author has applied worcs in several research projects:
	* https://github.com/cjvanlissa/storm_meta_father_mother
	* https://github.com/cjvanlissa/veni_sysrev
	* https://github.com/cjvanlissa/placebo_adhd
	* https://github.com/cjvanlissa/schumpe
	* https://github.com/cjvanlissa/coping_covid

@@

Response: We thank the Reviewer for this helpful suggestion! Although such user examples are not yet numerous, the lead author has six public WORCS repositories, and there are two public repositories by other users. In our opinion, it would be most useful to have a continuously cumulating list of example repositories, and link that list in the paper - rather than to publish a static list. Therefore, to address this comment, we have scripted a web scraper that searches GitHub for WORCS repositories (using the metadata tags created by the `worcs` package). We have embedded the scraper in the README.Rmd file on the worcs GitHub page, which is regularly updated. In the revised paper, we point readers to [this readme](https://github.com/cjvanlissa/worcs) for a list of user examples, in the Abstract, and when introducing the workflow.

@@

We agree that this is an important point, and have rewritten the paragraph comparing these two solutions (as cited below). We hope that this adresses the Reviewer's question. However, as the workflow of Peikert and Brandmaier is still in press, and the software implementation has not yet been fully developed - a more detailed comparison of these two approaches would be premature.

@@

\Assignment{CJ, AP, AB}
\WorkInProgress
\Medium

@@

We thank the Editor for pointing out that a full-blown user evaluation is not required for a resource paper.
Of course, we take this Reviewer suggestion very seriously, and intend to conduct a user evaluation after the paper is published and users begin to adopt the workflow.
In the meantime, we present the following evidence for WORCS' potential for reuse:

TODO

@@

To address the Reviewer's comment, we have made two changes: First, we now explicitly state, in the paper and the Workflow vignette, that the steps are not necessarily linear:  *"Note that, although the steps are numbered for reference purposes, we acknowledge that the process of conducting research is not always linear."*. Second, while it was always possible to manually add a manuscript or preregistration at a later point in time after creating a new project in RStudio - we have now added functions to make non-linear workflows even more convenient, specifically: `add_manuscript()` and `add_preregistration()`.

\Assignment{AP, AB}
\WorkInProgress
\Medium

>>>>

Dear Dr. Kuhn,

We would like to thank you for providing us with the opportunity to revise our manuscript, "WORCS: A Workflow for Open Reproducible Code in Science".
We sincerely appreciate the time you and the Reviewers have taken to read the manuscript, and the extensive and thought-provoking comments you have raised. 
We hereby submit our revision of the manuscript for your consideration. We have attempted to address each of the issues raised by yourself and the Reviewers on a point-by-point basis, and detail the changes made in the action letter below. We would like to thank you and the Reviewers for your time, and we look forward to receiving your response.

Sincerely, on behalve of all coauthors,
Caspar van Lissa

@@

To address the Reviewer's comment, we have set up an evaluation questionnaire, and sent it to the attendees of our previous workshops on WORCS. At the time of writing, responses are coming in. These are generally very favorable, and have not yet revealed major points of improvement.

We further want to point out that we have a strong history of addressing user comments submitted through the GitHub issues page.
WORCS adheres to the Core Infrastructure Initiative "best practices" (CII Best Practices),
which requires that
the channels through which users can provide feedback are clearly indicated,
as they are [here](https://github.com/cjvanlissa/worcs#contributing-and-contact-information).
It is also evident from the [resolved tickets on GitHub](https://github.com/cjvanlissa/worcs/issues?q=is%3Aissue+is%3Aclosed) that we have been very responsive in actively incorporating user feedback. This is another requirement of CII Best Practices.

We are committed to continue evaluating user experience, both actively through user evaluation forms, and passively through the GitHub issues pages.
We hope that this commitment, along with the evidence of our prior responsiveness to user feedback presented above, help assuage the Reviewer's concerns.

Finally, we wish to point out that the Editor has indicated that 'resource papers do not need to have a full-blown user evaluation, but we expect "sound evidence of its (potential for) reuse"'. We therefore also elaborate on the efforts we have made to ensure that WORCS is reusable, followed by evidence of its current adoption rates.

@@

There is also evidence that WORCS is already being adopted, even prior to publication (as of 14-11-2020):

1.  The preprint has been downloaded 582 times, since being published on
    31-05-2020
2.  The `worcs` R-package has been downloaded 2567 times from CRAN,
    since being published on 18-05-2020
3.  The GitHub project has been forked 8 times, watched 5 times, and
    starred 40 times
4.  The lead author has given invited lectures on WORCS at:
      - Max-Planck-Institute for Human Development (2020)  
      - Open Science Community Rotterdam (2020)  
      - Eindhoven University of Technology (2020)  
5.  WORCS is currently used in the following GitHub repositories (sorted
    by user):
      - cjvanlissa
          - [veni\_sysrev](https://github.com/cjvanlissa/veni_sysrev)  
          - [IHBT-SEM](https://github.com/cjvanlissa/IHBT-SEM)  
          - [schumpe](https://github.com/cjvanlissa/schumpe)  
          - [storm\_meta\_father\_mother](https://github.com/cjvanlissa/storm_meta_father_mother)  
          - [placebo\_adhd](https://github.com/cjvanlissa/placebo_adhd)  
          - [mpib](https://github.com/cjvanlissa/mpib)  
          - [coping\_covid](https://github.com/cjvanlissa/coping_covid)  
      - IndrajeetPatil
          - [worcs\_practice](https://github.com/IndrajeetPatil/worcs_practice)  
      - EstherPlomp
          - [RDM-survey](https://github.com/EstherPlomp/RDM-survey)  
      - lisallreiber
          - [worcs\_ws\_empty](https://github.com/lisallreiber/worcs_ws_empty)  
      - KimBaldry
          - [BIO-MATE](https://github.com/KimBaldry/BIO-MATE)

@@

Response: In addressing this comment, we respectfully refer to our response to the preceding comment ("There is no user evaluation", R1 C1), and again emphasize the Editor's comment that "resource papers do not need to have a full-blown user evaluation". We would also like to clarify that the goal of the present manuscript is to define the problem WORCS intends to address (specifically; meeting the TOP-guidelines for Open Science), introduce the tools WORCS uses to address this problem, explain how WORCS uses these tools, and provide a step-by-step workflow. It is not clear what "claims" are made that are unsupported by evidence. In the revision, we have attempted to further clarify this goal from the paper onset: REF.

We do take the suggestion of performing a user evaluation very seriously, but we also feel that it would be premature to do so even before the paper is published. We have built WORCS based on established tools and best practices, and we have gathered preliminary positive feedback from the community through workshops and teaching.

@@

Response: We thank the Reviewer for bringing this apparent contradiction to our attention.
The misunderstanding arises from a lack of explicit definitions of the terms "replication" and "reproducibility" in the original manuscript.
We are aware of the oft-lamented lack of consistent definitions for these terms (e.g., Patil, Peng, & Leek, 2019; Plesser, 2018).
In the revision, we have attempted to address this lack of clarity by including explicit definitions for these terms:

@@

After making these definitions explicit, it becomes clear that the Reviewer's definition of replication as "a necessary steps towards reproducibility" is not consistent with the definitions used in the present paper.

We further wish to clarify that the fact that we address guidelines 1-7, but not guideline 8 ("replication") is fully consistent with the intended purpose of the TOP-guidelines.
Specifically, the first seven guidelines are intended to ensure the "reproducibility of the reported results based on the originating data" [direct quote from @nosekPromotingOpenResearch2015a]. The eighth criterion, by contrast, is "not formally a transparency standard for authors, [but] addresses journal guidelines for consideration of independent replications for publication" (again a direct quote).
To avoid further confusion or controversy, we now address this distinction between the first 7 guidelines and the final 8th explicitly in the revised manuscript:

> These guidelines describe eight standards for open science,
the first seven of which "account for reproducibility of the reported results
based on the originating data, and for sharing sufficient information to conduct an independent 
replication" [@nosekPromotingOpenResearch2015a].
These guidelines are: 1) Comprehensive
citation of literature, data, materials, and methods; 2) sharing data, 3)
sharing the code required to reproduce analyses, 4) sharing new research
materials, and 5) sharing details of the design and analysis; 6)
pre-registration of studies before data collection, and 7) pre-registration of
the analysis plan prior to analysis.
The eighth criterion is "not formally a transparency standard for authors", but "addresses 
journal guidelines for consideration of independent replications for publication".
In this context, 
reproducibility can be defined as "re-performing the same analysis with the same code using a different analyst",
and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].
WORCS defines the goals of open science in terms of the first seven TOP-guidelines,
and the workflow and its `R` implementation are designed to facilitate compliance therewith.

@@

Response: We apologize; it appears that the reference list was not parsed correctly after rendering the manuscript to HTML - the preferred format of Data Science. The reference is indeed to the R-package. In the revised manuscript, we have repaired the references so that it is clear that this is a software package.

@@

The changes are extensive, so we will not quote them all here. Instead, we refer the Reviewer to the revised manuscript.

@@

\Assignment{BV, AB, AL, CJ}
\Done
\Medium

@@

Response: We agree completely that it is important to make open data findable beyond just providing it freely.
In order to address the Reviewer's comment, we've revised the paper and the workflow vignette to emphasize more clearly that an essential step of the workflow is to connect the GitHub repository to a project page on the Open Science Framework - a standard repository for open research.
We also devote more attention to the importance of DOIs; both for the project as a whole, and for specific resources, such as data files.
Finally, we have elaborated on the discussion of the FAIR principles, emphasizing that WORCS is not designed to address these comprehensively, but pointing out specific ways in which WORCS is consistent with these guiding principles. Please see our response to Reviewer 2, comment 2 for more details.

With regard to the specific platforms named by the Reviewer:
We are not aware of arguments to support a strong preference for Zenodo or FigShare over GitHub.
Without such arguments, as far as we are concerned, these repositories are largely interchangeable.
One key advantage of storing the data on GitHub is that they are then bundled with the
documentation (e.g., codebook), analysis code, and manuscript.
Thus, readers have access to **all** materials required for replication in one repository.
Moreover, the entire project can be readily updated from the console, and is version controlled.
GitHub additionally offers user interaction with the materials through
forking, cloning, opening issues, and sending pull requests -
all useful tools for replication and collaboration.

It is further worth noting that [GitHub is compatible with Zenodo](https://guides.github.com/activities/citable-code/);
users can document GitHub repositories in Zenodo and generate a DOI.
We now explicitly mention this option in the paper and workflow vignette.
We also mention the potential redundancy with Steps 15-16 of the workflow, however:
Connecting the GitHub repository to an OSF project and generating a DOI accomplishes effectively the same.

\Assignment{BV, AB, AL, CJ}
\Done
\Hard

@@

> We refer users of the `worcs` implementation in `R` to
[the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html),
which describes in more detail which steps are automated by the `R` package
and which actions must be taken by the user.
The vignette is also updated with future developments of the package.

@@

> We provide a conceptual outline of the WORCS procedure below, based on WORCS
Version 0.1.6.
The conceptual workflow recommends actions to take 
in the different phases of a research project, in order to work openly and transparently
in compliance with the TOP-guidelines and best practices for open science.
Note that, although the steps are numbered for reference purposes,
we acknowledge that the process of conducting research is not always linear.

@@

Response: We wish to emphasize that it was certainly unintended to suggest that everyone should us R for open science. To address this comment, we have done two things: First, throughout the paper, we have attempted to clarify the distinction between the platform independent conceptual workflow "WORCS", and its implementation for R users, `worcs`. Second, throughout the paper, we encourage the implementation of WORCS in other languages. Most illustrative of these changes is the start of the paragraph on **The R implementation of WORCS**:

@@

Response: We thank the Reviewer for these suggestions. Regarding the flowchart - we had already included one in the previous version of the manuscript. To address this comment, we have clarified the references in the text to this flowchart:

> The WORCS workflow constitutes a step-by-step procedure that researchers can follow to make a 
research project open and reproducible.
These steps are elaborated in greater detail below.
See Figure 1 for a flowchart that highlights important steps in different stages of the research process.

Regarding the example project, we have now curated a list of all public `worcs` projects on GitHub, and we reference this list in several places in the manuscript, including:

**Abstract**

> The source code for the `R` package and manuscript,
  and a list of [user examples of WORCS projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users), are available at 
  https://github.com/cjvanlissa/worcs.

**Introducing the workflow**

> For examples of how the workflow is used in practice,
we maintain [a list of public `worcs` projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users) on the WORCS GitHub page.

@@

Response: To address this comment, we have made three changes:

1) We have changed the workflow, emphasizing the importance of creating a project on the OSF and connecting it to the GitHub repository, and emphasizing the importance of DOIs and a license
2) We have rewritten the paragraph addressing the FAIR principles, with a more thorough discussion of how WORCS can help address each FAIR principle. We also state explicitly that *WORCS was not designed to comprehensively address these principles, but the workflow does facilitate meeting them.*
3) We have made minor changes to the `worcs` package, e.g., to include information on the project readme about how readers can obtain access to data.

This paragraph now reads:

> Aside from the TOP Guidelines, the FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a] are an important standard for open science.
These principles advocate that digital research objects should be Findable, Accessible, Interoperable and Reusable.
Initially mainly promoted as principles for data, they are increasingly applied as a standard for other types of research output as well,
most notably software [@lamprechtFAIRPrinciplesResearch2019].  
  WORCS was not designed to comprehensively address these principles, but the workflow does facilitate meeting them.
In terms of Findability, all files in GitHub projects are searchable by content and meta-data, both on the platform and through standard search engines.
The workflow further recommends that users create a project page on the Open Science Framework – a recognized repository in the social sciences.
When doing so, it is important to generate a Digital Object Identifier (DOI) for this project.
A DOI is a persistent way to identify and connect to an object on the internet, which is crucial for findability.
Additional DOIs can be generated through the OSF for specific resources, such as data sets.
It is further worth noting that, optionally, [GitHub repositories can be connected to Zenodo](https://guides.github.com/activities/citable-code/):
On Zenodo, users can store a snapshot of the repository, provide metadata, and generate a DOI for the project or specific resources.
When a project is connected to the OSF, as recommended, these steps may be redundant.
Finally, each worcs project has a [YAML file] (https://yaml.org/) that lists the data objects (and other meta-data) in the project, which will allow web crawlers to index worcs projects.
We use this metadata, for example, to identify public `worcs` projects on GitHub.  
  Accessibility is primarily safeguarded by hosting all project files in a public GitHub repository, which can be downloaded as a ZIP archive, and cloned or forked using the Git protocol. [GitHub is committed to long term accessibility of these resources]( https://archiveprogram.github.com/). We further encourage users to make their data accessible when possible.
When data must be closed, the `worcs` function `closed_data()` adds a paragraph to the project readme, indicating how readers can contact the author for access.  
  With regard to interoperability, the workflow recommends using plain-text files whenever possible, which ensures that code and meta-data are human- and machine-readable.
Furthermore, the `worcs` package is implemented in R, which is open source.
To promote reusability, we recommend users to specify an appropriate license for their projects.
The `worcs` package includes the CC licenses, and suggests a CC-BY 4.0 license by default.
This license is featured in the GitHub repository, and we encourage users to also display it on their project’s OSF page.
To further facilitate reusability, `worcs` automatically generates a codebook for
each data file.
We encourage users to elaborate on the default codebook by adding
variable description and categories, which would - in the future - enable
indexing data by topic area.

\Assignment{AL, BV, CJ}
\Done
\Medium

@@

Response: WORCS does not provide a service for de-identification,
and we consider this to be outside of the scope of the present paper.
If the Reviewer feels differently, we are open to being convinced, however!

The current team lacks relevant expertise to make this contribution, but
since `worcs` is open source software, anybody is welcome to contribute this functionality.
Such a contribution would warrant coauthorship on the package, in line with the contribution guidelines [outlined here](https://github.com/cjvanlissa/worcs#contributing-and-contact-information).
<!--What would this entail? Is it within scope?-->

@@

To address the Reviewer's comment, we have made two changes: First, we now explicitly state, in two places in the paper, and in the Workflow vignette included with the `worcs` package, that the steps are not necessarily linear:

**Introducing the workflow:**

> Note that, although the steps are numbered for reference purposes,
we acknowledge that the process of conducting research is not always linear.

**Limitations:**

> Another limitation is the fact that no single workflow can suit all research projects.
Not all projects conform to the steps outlined here;
some may skip steps, add steps, or follow similar steps in a different order.
In principle, WORCS does not enforce a linear order, and allows extentensive user flexibility.

Second, while it was always possible to manually add a manuscript or preregistration at a later point in time after creating a new project in RStudio - we have added more functions to increase the non-linear flexibility of the `worcs` package: `add_manuscript()` and `add_preregistration()`.

@@

Response: Respectfully, we are not entirely sure what the criticism is here.
First, as addressed in our response to the previous comment, there seems to be a misunderstanding about the scope of the paper. To clarify, we again state that our paper introduces a workflow for open science, based on objective criteria set out in the TOP-guidelines. Regarding novelty: Aside from the work-in-press of our co-authors - which is focused on strict computational reproducibility and is compatible with the present workflow - this is one of the first efforts to translate the requirements of open science into a step-by-step workflow. When you search Google Scholar for "open science workflow", the first hit that actually describes a workflow is the present manuscript.

@@

> Git tracks changes to files on a line-by-line basis.
The user can store these changes by making a "commit": a 
snapshot of the version controlled files.
Each "commit" can contain as many changes as desired;
for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
Each commit is further tagged with a message, describing the changes.
The `R` implementation of `worcs` contains a convenience function,
`git_update("your commit message")`,
which creates a new commit for all changes since the previous commit,
labels it with "your commit message", and pushes these changes to the remote repository.
It is, effectively, a Git "quick save" command.

@@

Response: We thank the Reviewer for acknowledging the idea to be interesting.
Whether it will be adopted by researchers remains to be seen after publication.
We can offer some prelimiary evidence that the workflow is already being adopted (as of 14-11-2020):

1.  The preprint has been downloaded 582 times, since being published on
    31-05-2020
2.  The `worcs` R-package has been downloaded 2567 times from CRAN,
    since being published on 18-05-2020
3.  The GitHub project has been forked 8 times, watched 5 times, and
    starred 40 times
4.  The lead author has given invited lectures on WORCS at:
      - Max-Planck-Institute for Human Development (2020)  
      - Open Science Community Rotterdam (2020)  
      - Eindhoven University of Technology (2020)  
5.  WORCS is currently used in the following GitHub repositories (sorted
    by user):
      - cjvanlissa
          - [veni\_sysrev](https://github.com/cjvanlissa/veni_sysrev)  
          - [IHBT-SEM](https://github.com/cjvanlissa/IHBT-SEM)  
          - [schumpe](https://github.com/cjvanlissa/schumpe)  
          - [storm\_meta\_father\_mother](https://github.com/cjvanlissa/storm_meta_father_mother)  
          - [placebo\_adhd](https://github.com/cjvanlissa/placebo_adhd)  
          - [mpib](https://github.com/cjvanlissa/mpib)  
          - [coping\_covid](https://github.com/cjvanlissa/coping_covid)  
      - IndrajeetPatil
          - [worcs\_practice](https://github.com/IndrajeetPatil/worcs_practice)  
      - EstherPlomp
          - [RDM-survey](https://github.com/EstherPlomp/RDM-survey)  
      - lisallreiber
          - [worcs\_ws\_empty](https://github.com/lisallreiber/worcs_ws_empty)  
      - KimBaldry
          - [BIO-MATE](https://github.com/KimBaldry/BIO-MATE)

@@

Response: We thank the Reviewer for this helpful suggestion! Although such user examples are not yet numerous, the lead author has six public WORCS repositories, and there are four public repositories by other users. In our opinion, it would be most useful to have a continuously cumulating list of example repositories, and link that list in the paper - rather than to publish a static list. Therefore, to address this comment, we have scripted a web scraper that searches GitHub for WORCS repositories (using the metadata tags created by the `worcs` package). We have embedded the scraper in the README.Rmd file on the worcs GitHub page, which is regularly updated. In the revised paper, we point readers to [a list of public `worcs` projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users) in the Abstract, and when introducing the workflow.

@@

We agree that this is an important point. We have attemptedto address it by rewriting the paragraph comparing these two solutions (as cited below). We hope that this is satisfactory. At this point in time, this is all we can say about the comparison of the two workflows, because the workflow of Peikert and Brandmaier is still in press, and the software implementation has not yet been fully developed.

@@

\Assignment{CJ, AP, AB}
\Done
\Medium

@@

We thank the Editor for pointing out that a full-blown user evaluation is not required for a resource paper.
Of course, we take this Reviewer suggestion very seriously, and intend to conduct a user evaluation after the paper is published and users begin to adopt the workflow.
Please refer to our response to Reviewer 1's comment 1, where we discuss our efforts to ensure the *potential for reuse*, and provide preliminary evidence for actual reuse/uptake of the workflow.

@@

To address Reviewer 3's comment, we have made two changes: First, we now explicitly state, in two places in the paper, and in the Workflow vignette included with the `worcs` package, that the steps are not necessarily linear:

**Introducing the workflow:**

> Note that, although the steps are numbered for reference purposes,
we acknowledge that the process of conducting research is not always linear.

**Limitations:**

> Another limitation is the fact that no single workflow can suit all research projects.
Not all projects conform to the steps outlined here;
some may skip steps, add steps, or follow similar steps in a different order.
In principle, WORCS does not enforce a linear order, and allows extentensive user flexibility.

Second, while it was always possible to manually add a manuscript or preregistration at a later point in time after creating a new project in RStudio - we have added more functions to increase the non-linear flexibility of the `worcs` package: `add_manuscript()` and `add_preregistration()`.
